---
title: "Inverse Oxygen Model"
author: "Alice Carter"
date: "6/24/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dygraphs)
library(xts)
library(lubridate)
source("~/../git/oxygen_proc_model/src/batch_nhd_retrevial.R")
source("~/../git/oxygen_proc_model/src/oxygen_dynamics_functions.R")
```

## Oxygen process model
Start with a mass balance equation for oxygen concentration C:

$$\frac{\delta VC}{\delta t} = -\left(\frac{\delta q_x}{\delta x} + \frac{\delta q_y}{\delta y} + \frac{\delta q_z}{\delta z}\right) + S$$

where:
$V$ = volume
$C$ = concentration
$q$ = mass fluxes due to advective and dispersive transport
$S$ = sinks and sources

Assume:
Uniform flow/well mixed water column:

$$\frac{\delta q_x}{\delta x} + \frac{\delta q_y}{\delta y} = 0$$
$$\frac{\delta VC}{\delta t} = -\frac{\delta q_z}{\delta z} + S$$

Integrate with respect to z and define fluxes per unit ground area A assuming the channel is rectangular with a depth of D

$$\frac{\delta H\overline{C}}{\delta t} = q(H) - q(0) + \int_0^H Sdz$$

where:
$\overline{C}$ is the depth-averaged concentration,
$H$ is the depth 
$q(H) = k(C_{sat} - \overline{C})$ is the exchange of oxygen with the atmosphere 
$B(t) = -q(0) + \int_0^DSdz$ represents biological production and consumption of oxygen 

So we can describe the biological time series as a function of $\overline{C}(t)$, depth, discharge, and riverbed slope using empirical equations:

$$ B(t) = \frac{d(H\overline{C})}{dt} - k(C_{sat} - \overline{C})$$

## Example Stream: 
# New Hope Creek

```{r, include = FALSE}
dat <- read_csv("C:/Users/Alice Carter/git/nhc_50yl/NHC_2019_metabolism/data/metabolism/processed/NHC.csv",
                guess_max = 10000)
dat <- filter(dat, year(DateTime_UTC) == 2019)

# load site data
site <- read_csv("C:/Users/Alice Carter/git/nhc_50yl/NHC_2019_metabolism/data/siteData/NHCsite_metadata.csv")[1,] %>%
  select(sitecode, latitude, longitude, CRS, slope)
nhc_slope <- site$slope
# xts(dat$DO.obs, dat[,1, drop = T]) %>%
#   dygraph(ylab = "DO (mg/L)", main = "NHC dissolved oxygen")
```

```{r, echo = FALSE}
dat %>% 
  filter(month(dat$DateTime_UTC) == 3) %>%
  ggplot(aes(DateTime_UTC, DO.obs)) +
  geom_line() 
```

```{r}
m_dat <- dat %>%
  # filter(!is.na(DO.obs)) %>%
  # select(DateTime_UTC, DO.obs, avg_velocity, DO.sat, depth, temp.water) %>%
  mutate(slope = nhc_slope,
         # v_ms = calc_vel_from_disch(discharge),
         k600_mean = calc_k600_raymond5(slope, avg_velocity), # m/d
         # k600_min = calc_k600_raymond5(slope, avg_velocity)[2],
         # k600_max = calc_k600_raymond5(slope, avg_velocity)[3],
         k_O2_mean = kO2fromk600(k600_mean, temp.water),
         q_D_gm2d = k_O2_mean * (DO.sat - DO.obs), # m/d * mg/L  = g/m2d
         dHCdt_gm2d = c(NA, diff(dat$DO.obs))*depth/15*60*24, # mg/L/min * 60 min/h * 24 h/d
         B_gm2d = dHCdt_gm2d - q_D_gm2d) # biological timeseries mg/m2/d
         
m_dat %>%
  filter(month(DateTime_UTC) == 3) %>%
  ggplot(aes(DateTime_UTC, dHCdt_gm2d)) +
  geom_line() +
  geom_line(aes(DateTime_UTC, B_gm2d), col = 2)
```

```{r}
daily <- m_dat %>%
  mutate(date = as.Date(with_tz(DateTime_UTC, tz = "EST"))) %>%
  group_by(date) %>%
  summarize_all(list(mean = mean, min = min, max = max)) %>%
  select(date, DO_mean = DO.obs_mean, starts_with(c("B_")), 
         depth_m = depth_mean, q_m3s = discharge_mean, temp_C = temp.water_mean,
         spc_mScm = SpecCond_mScm_mean, turb = Turbidity_mV_mean, 
        light = light_mean)
pairs(daily)
```
